---
title: "Course project, Regression Models"
output: pdf_document
---

# explore the mtcars dataset

First here are the brief explanations for all variables. 
```
?mtcars
mpg 	Miles/(US) gallon
cyl 	Number of cylinders
disp 	Displacement (cu.in.)
hp 	  Gross horsepower
drat 	Rear axle ratio
wt 	  Weight (lb/1000)
qsec 	1/4 mile time
vs 	  V/S
am 	  Transmission (0 = automatic, 1 = manual)
gear 	Number of forward gears
carb 	Number of carburetors 
```

```{r}
data(mtcars)
# convert categorical variables into factor
mtcars$am = factor(mtcars$am)
mtcars$cyl = factor(mtcars$cyl)
mtcars$vs = factor(mtcars$vs)
# summary(mtcars)
# variable selection
summary(lm(mpg ~ . , data = mtcars))$coefficient
fit <- lm(mpg ~ . , data = mtcars)
# calculate the variance inflation factor (VIF)
library(car)
sqrt(vif(fit))
```
It seems that the weight is the best predictor for mpg. Therefore, I will include only weight and am as the covariates in my following analysis.

# model fitting and selection
```{r}
# to select the best model, I'll fit 3 models and compare them
fit1 = lm(mpg ~ am, data = mtcars)
summary(fit1)$coefficient
fit2 = lm(mpg ~ wt + am, data = mtcars)
summary(fit2)$coefficient
fit3 = lm(mpg ~ wt * am, data = mtcars)
summary(fit3)$coefficient
anova(fit1, fit2, fit3)
```

From model fit1, I found that manual transmission has positive effect on mpg comparing with automatic transmission. The coefficient 7.2 means the manual transmission has 7.2 mpg increase than automatic transmission on average.      
From model fit3, I also found that the interaction between weight and transmission is significant. The intercept 31.4 is the average mpg of cars from the reference group am 0 (automatic transmission) at weight 0. -3.8 is the change of mpg for each 1000 pounds from the reference group am 0. 14.9 is the increase of mpg comparing am1 cars with am0 cars at weight 0. 

```{r fig.width=6.5, fig.height=4}
# to show that there is interaction, I plotted the regression within each am group
library(ggplot2)
theme_set(theme_bw())
qplot(wt, mpg, data = mtcars, geom="point", colour = am) + geom_smooth(method="lm")
```

I can also observe the intersection of regressin line, suggesting the interaction between weight and transmission. The effect of weight on the mpg is dependant on the transmission type of cars. That is why the slopes of two regression lines are different.

# residues and diagnostics
```{r fig.width = 4, fig.height = 4}
plot(predict(fit3), resid(fit3))
```
```{r fig.height = 5.5}
par(mfrow = c(2, 2))
plot(fit3)
# round(dfbetas(fit3)[, 2], 3)
round(hatvalues(fit3), 3)
```
There is no abnormal pattern found in the residues plot.

# executive summary
1. There is significant evidence suggesting that an manual transmission is better for MPG than an automatic transmission.
2. The MPG difference between automatic and manual transmissions is 7.2 mpg disregarding the weight, while the difference is 14.9 mpg regarding the weight.
